\section{General Response}

We thank the Associate Editor and all five reviewers for their careful and
constructive evaluation. The reviews were detailed and fair, and they have
substantially improved the manuscript. We are encouraged that the reviewers
found the problem \textbf{important} (R1, R2, R4), the functional-decoupling
perspective \textbf{theoretically well-motivated} (R5), and the framework
\textbf{architecturally clean, technically sound, and empirically interesting} --
including gains over a strong CLIP ViT-L/14 baseline and larger improvements on
hard intent categories (R1, R5).

We have grouped the major concerns into five themes and addressed each one with
new experiments, new analysis, and substantial rewriting:

\begin{itemize}
  \item \textbf{Fairness and interpretability of the comparisons; the effect of
  class-wise thresholding (R1).} We added a controlled protocol in which all
  in-house baselines and FDIL share the \emph{same} CLIP ViT-L/14 features, the
  \emph{same} train/val/test split, and the \emph{same} validation-only
  thresholding protocol, together with a four-cell grid that decomposes the
  contribution of the \emph{method} from that of \emph{calibration} (global vs.\
  class-wise thresholding), reported as mean$\pm$std over 3 seeds. FDIL improves
  over the CLIP baseline under \emph{both} thresholding protocols, and the
  threshold-free mAP confirms the gain is not a calibration artifact. We further
  re-implemented representative external methods (HLEG, LabCR, IntCLIP) over the
  \emph{same} frozen CLIP ViT-L/14 features: once the backbone is equalized they
  fall to the level of the plain CLIP baseline, while FDIL still leads them by
  $+3.4$ to $+4.7$ average F1, confirming the improvement is algorithmic rather
  than backbone-driven.

  \item \textbf{Novelty and theoretical depth (R2, R4, R1).} We added a formal,
  information-theoretic / feature-selection motivation for \emph{why} the two
  ambiguities are most naturally handled by orthogonal components, a head-to-head
  unified-vs-decoupled comparison, negative-control experiments showing that
  arbitrary VLM text priors do \emph{not} reproduce the gains, and a controlled
  generalization study on a second (affect) dataset on which privileged
  distillation significantly beats the baseline where its precondition holds.

  \item \textbf{Presentation and clarity (R4, R5, R3).} We rewrote the
  Introduction (clear purpose, critical analysis of how existing methods fail on
  each ambiguity type, expanded LLM-based intent-understanding survey), added a
  notation table and a Discussion section, fixed the Figure~2 caption, corrected
  the ``FDIR''$\rightarrow$``FDIL'' typos, and proofread the manuscript and
  bibliography.

  \item \textbf{Stronger and larger-scale evaluation (R4, R3).} We added a
  controlled cross-validation generalization study on a second (affect) dataset
  with paired significance testing, comparisons
  against 2025/2026 large multimodal models, t-SNE embedding
  visualizations, ROC/PR curves, threshold-free metrics with statistical
  significance testing, and an extended efficiency comparison
  (params/FLOPs/latency). In the new zero-shot MLLM comparison, GPT-4o reaches
  $51.52$ average F1 on the test split, while Qwen3-VL-8B and InternVL3-8B
  reach $39.63$ and $39.97$, respectively; all remain below the trained
  CLIP/FDIL models, clarifying that prompting a general MLLM alone does not
  solve Intentonomy.

  \item \textbf{Reproducibility (R1, EiC).} We documented the validation-only
  selection protocol, expanded implementation details, and replaced the
  ``available on request'' statement with a concrete artifact-release plan.
\end{itemize}

\paragraph{Note on the page limit.} We retained all reviewer- and editor-requested
analyses in the main manuscript, but substantially condensed their presentation to
comply with the Pattern Recognition 35-page limit. Specifically, we (i) moved the
notation table into the Method section and reduced it to the core symbols, (ii)
merged the t-SNE visualization and per-class attribution into a single compact
interpretability subsection, (iii) merged the gating and teacher-modality
ablations into one table, and (iv) shortened the related-work, theoretical-derivation,
and discussion text. No analysis or result was removed: where a table was condensed
out of the main text, its full version is reproduced in this letter. In particular,
the complete per-class F1 attribution is given in Table~\ref{tab:per_class_attr} of
our response to Reviewer~1.

Below we provide point-by-point responses to each reviewer. Newly added or
revised content is referenced by section, table, or figure. In the revised
manuscript, all changes are highlighted in \revised{red}.
