% ================================
% === Reviewer 2 ===
% ================================
\reviewerblock{Reviewer~2}{%
We thank the reviewer for acknowledging that the manuscript is well-written, technically sound, and empirically validated, and that distinguishing semantic from supervisory uncertainty is appealing. We take the novelty concern very seriously and have addressed it on three fronts so that the contribution is a \emph{principle}, not a combination of existing components.}

\cresp{The major drawback is insufficient novelty: the idea of distinguishing semantic uncertainty from supervision uncertainty is appealing but does not constitute a new theoretical principle or algorithmic idea. The results appear to be based mainly on the correct combination of existing methods rather than the development of a novel approach.}
{We address this on four fronts.

\smallskip\noindent\textbf{(a) A formal decoupling principle (new theory, \S\ref{sec:decoupling_principle}).} We now formalize the two ambiguities as acting on \emph{different terms of the learning objective} (Eq.~\eqref{eq:risk_decomp}). The starting point is an empirical diagnostic that the core failure is often ranking/calibration under local ambiguity, not global semantic retrieval: the visual baseline's Top-10 candidates already cover \res{$83.9\%$} of positive labels, \res{$95.1\%$} of test images contain at least one ground-truth intent in that Top-10 set, and \res{$67.8\%$} of baseline false negatives are still recoverable from this same candidate set. An oracle restricted to the Top-10 neighborhood would reach \res{$87.4/91.3/89.7$} Macro/Micro/Samples F1, far above the realized baseline. This is why the new theory models semantic ambiguity as local feature-side aliasing among retrieved neighbors rather than as simple absence of the correct class. Formally, supervisory ambiguity is a label-side conditional-entropy term that depends only on $\Pr(Y\mid X)$ and is irreducible by any function of the features, whereas semantic ambiguity is a feature-side inter-class mutual-information aliasing term (the semantic term~(F) of Eq.~\eqref{eq:risk_decomp}) that depends only on the representation $\phi$. Because the two are functionals of different objects, we show their corrections have non-interfering first-order effects: a reliability gate that is a function of the agreement $\omega$ alone (UTD, label side) and a reranking/residual that is a function of $(\phi,\text{class text})$ alone (SLR-C, feature side) act on disjoint factors of the risk, so a single scalar-weighted mechanism that mixes both signals must trade one ambiguity against the other while a decoupled realization need not. This reframes the contribution as a principle with three testable predictions (separable additive threshold-free increments; content-dependence of the teacher; non-uniqueness of the realization), each verified in Section~\ref{sec:experiments}. We are careful not to overclaim necessity: as prediction~(P3) and the head-to-head comparison in~(b) show, a unified control built on a single ambiguity signal comes within about one average-F1 point of the decoupled model, while naively fusing both signals into one target underperforms it, so we frame decoupling as the more interpretable, separately measurable, and independently deployable realization of the principle -- and a modestly more accurate one -- rather than the only route to a competitive accuracy.

\smallskip\noindent\textbf{(b) Unified vs.\ decoupled, head to head (new table).} We added a controlled comparison in which a single unified mechanism is given the same inputs and budget as decoupled FDIL. A unified control that mixes SLR-C probabilities and rationale-teacher probabilities into one joint distillation target -- removing residual decomposition and agreement-aware gating -- reaches only $55.66$ average F1 (3-seed mean), \emph{below} decoupled FDIL ($57.49$; Macro $50.69\!\to\!52.13$, Hard $32.41\!\to\!35.02$, mAP $54.60\!\to\!55.10$) and at the level of the weaker single signal. The strongest single-signal control (SLR-C only) is closer, $0.62$ average F1 behind FDIL ($56.87$ vs.\ $57.49$) and marginally higher on mAP. We therefore state the claim as follows: naively fusing the two ambiguity signals into one supervision target is suboptimal, and decoupling yields a modest but consistent aggregate-accuracy benefit on top of its primary advantages -- interpretability, separate measurability, and independent deployability.

\smallskip\noindent\textbf{(c) Generalization to a second, larger dataset.} We have substantially strengthened the EMOTIC probe, and it now provides positive evidence of generalization rather than an inconclusive ``no transfer'' report: once its precondition---per-sample annotator agreement at training time---is satisfied, the privileged-distillation framework carries over to this affect domain and improves the image-only baseline by \emph{statistically significant} margins. Two observations organize the finding.

First, the earlier flat result on EMOTIC's standard split reflects a property of that benchmark's annotation protocol rather than a shortcoming of the mechanism. Our agreement-aware gating reweights the rationale-teacher term by per-sample annotator agreement, but EMOTIC multi-annotates only its validation and test splits; the training split is single-annotator, so every training instance carries a degenerate agreement of unity and the gated term reduces to standard knowledge distillation by construction. We verified that the gated and ungated students are numerically identical on this split. EMOTIC's affect labels are, moreover, already well predicted by the CLIP baseline, which leaves little headroom on the standard split.

Second, when the framework is evaluated in the regime for which it was designed, distillation yields clear and significant gains. We pool EMOTIC's multi-annotated validation and test instances ($10{,}613$ persons, each carrying a fractional agreement label) and perform $5\times5$ stratified cross-validation entirely within this pool, so that label construction is consistent across folds and the gate is active on \res{$97.7\%$} of samples (the mean gate weight rising from $0$ to $0.51$). Privileged distillation then improves the image-only baseline on every metric---mAP \res{$42.57\!\to\!42.77$} ($+0.20$), Macro \res{$+0.39$}, Micro \res{$+0.65$}, and Samples-F1 \res{$+0.67$}---with each improvement paired-significant over the $25$ seed$\times$fold runs ($p<0.01$; Table~\ref{tab:emotic_pool}). We also observed that in-sample teacher targets are overfit ($87$ versus $66$ mAP), and that adopting out-of-fold teacher predictions is what prevents distillation from degrading the student; we therefore use out-of-fold targets throughout.

Once agreement is present, the gate is no longer inert. It activates on nearly all samples, and, consistent with its intended mechanism, its benefit concentrates on the low-agreement instances it is designed to protect. In aggregate over the $25$ runs, however, the gated variant is on par with standard distillation rather than significantly above it (all gate-versus-KD contrasts within error bars, $|\Delta|\!\le\!0.01$~mAP, $p>0.6$), which we attribute to EMOTIC's agreement being borrowed from evaluation annotators and therefore coarser than Intentonomy's native three-annotator training-time signal. We accordingly present EMOTIC as evidence of \emph{framework-level} generalization---privileged distillation transfers to a second, affect domain and significantly outperforms the baseline, and the gate behaves as designed wherever agreement is available---while locating the gate's decisive incremental advantage on Intentonomy, whose rich training-time agreement makes it identifiable. The core novelty argument rests on the formal principle in~(a), the matched-backbone gains, the unified-versus-decoupled comparison in~(b), and the negative controls below, and we state this scope explicitly in the Limitations (Section~\ref{sec:discussion}).

\begin{table}[t]\centering\small
\caption{Controlled $5\times5$ cross-validation on the multi-annotated EMOTIC
val+test pool (per-sample agreement available at train time; mean over $25$
seed$\times$fold runs). Privileged distillation significantly improves the
image-only baseline. With agreement now available the gate activates and
concentrates its benefit on low-agreement samples, while in aggregate it matches
standard KD on this pool, where the borrowed agreement signal is coarser than
Intentonomy's native training-time agreement.
$^{*}p<0.01$, paired $t$-test vs.\ baseline over the $25$ runs.}
\label{tab:emotic_pool}
\begin{tabular}{lcccc}
\hline
Method & mAP & Macro-F1 & Micro-F1 & Samples-F1 \\
\hline
Image-only baseline            & $42.57$ & $43.34$ & $60.32$ & $60.01$ \\
\;\;+ standard KD              & $42.77^{*}$ & $43.68^{*}$ & $60.95^{*}$ & $60.75^{*}$ \\
\;\;+ agreement-gated KD       & $42.77^{*}$ & $43.72^{*}$ & $60.97^{*}$ & $60.67^{*}$ \\
\hline
Oracle teacher (privileged text) & $66.43$ & $60.96$ & $74.63$ & $75.68$ \\
\hline
\end{tabular}
\end{table}

\smallskip\noindent\textbf{(d) The two modules contribute separately and additively (threshold-free).} On mAP, UTD alone contributes \res{$+2.40$} over the CLIP baseline ($50.67\!\to\!53.07$) and adding SLR-C contributes a further \res{$+2.03$} ($53.07\!\to\!55.10$), for a total of \res{$+4.43$}. Because each decoupled module yields an independent, measurable increment on a threshold-free metric, the decoupling is empirically separable rather than a single re-labeled mechanism.}

\cresp{The VLM / text prior may be mere external-knowledge stacking.}
{We added negative-control retrains (new Table~\ref{tab:negative_controls}) that corrupt one ingredient of the full pipeline at a time. The decisive control is the \emph{shuffled rationale}, where each image keeps a teacher rationale but it belongs to a \emph{different} image. This is the single most damaging change we tried -- it costs \res{$7.8$ Macro F1} and \res{$7.6$ Hard F1}, i.e.\ \emph{worse than removing the teacher altogether} (w/o UTD) -- because a mismatched rationale actively contradicts the image rather than merely withholding information. This directly rules out the ``any external text'' explanation: the gain depends on the per-image \emph{content} of the rationale, not on the mere presence of additional text or a calibration effect. Removing the uncertainty gate (ungated) or flattening it to a constant (uniform gate) costs a smaller, more mixed amount, so the agreement gate is a modest refinement rather than the main source of value. These controls focus on the uncertainty-aware rationale teacher because it is the load-bearing component of UTD, which is precisely the part R2 worried might be generic external knowledge.}
