% ================================
% === Reviewer 3 (Associate Editor metareview) ===
% ================================
\reviewerblock{Reviewer~3 (Associate Editor)}{%
We thank the Associate Editor for the focused, actionable points.}

\cresp{The authors are encouraged to clearly state, in the Abstract, the quantitative improvements achieved by the proposed FDIL method compared to existing approaches.}
{Following this recommendation, we have revised the Abstract to report the controlled gains over representative recent methods: at least \res{$+3.4$ Avg F1}, \res{$+5.7$ Hard F1}, and \res{$+3.4$ threshold-free mAP} under matched CLIP ViT-L/14 features and validation-only thresholding.
\revloc{Abstract (final sentence) and Table~\ref{tab:main_results}}
\revshot{abstract}}

\cresp{In Section~3.3.1, Semantic Prior Construction, the authors should provide a more detailed justification for the selected scenarios, including their representativeness and the expected outcomes in conditions not covered by the presented scenarios.}
{We appreciate the request for an explicit selection criterion and uncovered-case behavior. The revised Semantic Prior Construction now clarifies that the scenarios are not hand-picked. A fixed generation protocol is applied uniformly to every intent class and enumerates visually distinct realizations along five axes---subjects/objects, actions, settings, atmosphere, and photographic style---with at least three settings per intent to span intra-class variation rather than a single canonical view. The objective is representative coverage, not an exhaustive list of every possible scene. Each source is normalized independently, and the scenario score is averaged with lexical and canonical scores before being used only to rerank the base model's Top-$K$ candidates. Consequently, an image outside the generated scenarios does not trigger a brittle lookup failure: the scenario match weakens, while the lexical and canonical sources, unchanged base logits, and image-conditioned residual student continue to support the prediction. The prior-source ablation further shows that the three sources are complementary and that their ensemble provides the best aggregate coverage.
\revloc{Sec.~\ref{sec:method} (Semantic Prior Construction) and Table~\ref{tab:prior_ablation}}
\revshot{scenario_prior}}

\cresp{Table~2 shows that FDIL achieves lower performance than the conventional HLEG method on the Easy subset, yet this observation is not discussed. The authors should explain the possible reasons for this result and clarify whether certain characteristics of the Easy subset contribute to the similar performance between the two methods. Further discussion of the key observations in the table is recommended.}
{This Easy-subset exception is now stated explicitly rather than absorbed into the aggregate result. FDIL scores \res{$81.19$}, only $0.30$ below HLEG and slightly above the CLIP baseline ($80.99$). As discussed in the revised manuscript, Easy classes are largely visually unambiguous; a strong CLIP representation already separates them well, leaving little headroom for modules designed to resolve semantic overlap and uncertain supervision. The more diagnostic pattern appears on the Medium and Hard subsets: FDIL reaches \res{$57.73$} and \res{$35.02$}, respectively, and its gains over the CLIP baseline are substantially larger there. Per-class attribution is consistent with this explanation: the largest improvements occur in abstract Hard classes such as \textit{WorkILike}, \textit{PassionAboutSomething}, and \textit{SocialLifeFriendship}, while only one Easy class shows a small regression ($-0.96$ F1). We therefore present the near-tie on Easy as an expected scope characteristic of an ambiguity-targeted method, not as evidence of uniform dominance.
\revloc{Sec.~\ref{sec:experiments} (Performance on Difficulty- and Content-Dependent Subsets), Table~\ref{tab:difficulty_results}; supporting per-class values in Table~\ref{tab:per_class_attr} of this letter}
\revshot{difficulty_results}}
