% ================================
% === Reviewer 4 ===
% ================================
\reviewerblock{Reviewer~4}{%
We thank the reviewer for the thorough review, which we address point by point.}

\cresp{The manuscript must require careful proofreading, such as [7] IEEE Trans.\ Affect.\ Comput., and so on.}
{Thank you for catching this presentation issue. We performed a manuscript-wide proofreading pass and standardized bibliography venue names and abbreviations, including the flagged \emph{IEEE Trans.\ Affect.\ Comput.} entry.
\revloc{Bibliography and manuscript-wide language}
\revshot{proofreading}}

\cresp{The paper is not very clear but only claims that `` A central difficulty of this task lies in two complementary forms of ambiguity: supervisory ambiguity, caused by subjective annotation over plausible intentions, and semantic ambiguity, caused by semantically adjacent and overlapping intents.'' Please give more analysis about these disadvantages in the existing visual intention understanding.}
{We appreciate this request for a sharper diagnosis. The Introduction and Related Work now explain that structure-aware and semantic-prior methods mainly address class-side overlap, whereas robust-loss and disagreement-aware methods mainly address supervision-side uncertainty; treating either as the sole source leaves the other failure mode unresolved. % 加一下，在method里我们给了empirical motivation和理论解释，说明我们的方法是有理论依据的
\revloc{Sec.~\ref{sec:intro} and Sec.~\ref{sec:related_work}}
\revshot{ambiguity_analysis}}

\cresp{The clear purpose of work the proposed work is lacking in the introduction section.}
{We have rewritten the Introduction to state the purpose directly: FDIL converts external semantic knowledge into a structured prior and an uncertainty-gated training signal that target semantic and supervisory ambiguity separately.
\revloc{Sec.~\ref{sec:intro} (final related-work transition and contribution paragraphs)}
\revshot{purpose}}

\cresp{The introduction should be re-discussed and extended. I think that more Large Language model-based visual intention understanding methods should be mentioned. In the introduction, this section is hard for readers to understand its latest development.}
{To bring the literature discussion up to date, we have expanded it to include IntCLIP (ECCV~2024), IntentMLM (Pattern Recognition~2026), and the broader use of MLLMs as external semantic sources. Specifically, the revised text distinguishes semantic knowledge supplied by these models from the dataset-specific calibration and uncertainty-aware supervision required by FDIL, and links this distinction to the new zero-shot MLLM comparison.
\revloc{Sec.~\ref{sec:intro}, Sec.~\ref{sec:related_work}, and Table~\ref{tab:main_results} (Zero-shot VLM group)}
\revshot{recent_mllm_work}}

\cresp{You may consider including a table of notations.}
{We have followed this suggestion and added a compact notation table covering the symbols used in the method and theoretical analysis. % 加一句由于page limit我们列了主要符号
\revloc{Sec.~\ref{sec:method}, Table~\ref{tab:notation}}
\revshot{notation}}

\cresp{There should be larger-scale datasets to better evaluate the proposed methods.}
{We share the reviewer's view that evidence beyond Intentonomy is useful for assessing scope. We have added a controlled $5\times5$ cross-validation study on the multi-annotated EMOTIC pool, using out-of-fold teacher predictions so that agreement-aware distillation is evaluated without target leakage. Specifically, standard and agreement-gated distillation each improve the image-only baseline on all four metrics ($p<0.01$ over 25 runs), while the gated variant is comparable to standard distillation in aggregate. We therefore use EMOTIC as secondary evidence of framework-level transfer and retain Intentonomy as the dataset that establishes the incremental value of native training-time agreement. % 写得更详细一点，当有agreement信息时，gated的效果更好。这里应该还有一张表，展现不同agreement下的结果，其中agreement较低时效果更好。之前跑过结果，需要找一下并绘制表格。加几句说我们把Emotic的结果和分析写进了discussion里，说明了我们的方法在Emotic上也有一定的泛化能力

\begin{table}[t]\centering\small
\caption{Secondary generalization evidence on the multi-annotated EMOTIC validation and test pool: controlled $5\times5$ cross-validation, mean over $25$
seed$\times$fold runs. $^{*}p<0.01$, paired $t$-test vs.\ baseline.}
\label{tab:emotic_pool}
\begin{tabular}{lcccc}
\hline
Method & mAP & Macro-F1 & Micro-F1 & Samples-F1 \\
\hline
Image-only baseline            & $42.57$ & $43.34$ & $60.32$ & $60.01$ \\
\;\;+ standard KD              & $42.77^{*}$ & $43.68^{*}$ & $60.95^{*}$ & $60.75^{*}$ \\
\;\;+ agreement-gated KD       & $42.77^{*}$ & $43.72^{*}$ & $60.97^{*}$ & $60.67^{*}$ \\
\hline
Oracle teacher (privileged text) & $66.43$ & $60.96$ & $74.63$ & $75.68$ \\
\hline
\end{tabular}
\end{table}

\revloc{Sec.~\ref{sec:discussion} (Limitations and Future Directions); auxiliary results in Table~\ref{tab:emotic_pool} of this letter}
\revshot{discussion_scope}}

\cresp{Experimental evaluation is not sufficient. More comparisons with state-of-the-art Large Language Model methods (such as 2025 and 2026) and t-SNE are needed.}
{In response, we have added zero-shot Qwen3-VL-8B-Instruct and InternVL3-8B under the same label set and output protocol, together with the published GPT-4o result from IntentMLM. Specifically, all three prompt-only models remain below the task-trained CLIP/FDIL models, so we treat MLLMs as useful semantic sources rather than substitutes for task-specific learning. We also added a t-SNE comparison on semantically adjacent intents, with the complete scores retained in the main-results table.
\revloc{Sec.~\ref{sec:experiments} (Zero-shot Multimodal Large Model Comparison and Interpretability Analysis), Table~\ref{tab:main_results}, and Fig.~\ref{fig:tsne}}
\revshot{zero_shot_and_tsne}}

\cresp{Please use some ROC-based curves for a clear evaluation of the proposed method with other state-of-the-art methods.}
{We have added micro-averaged ROC and precision--recall curves for FDIL, the CLIP baseline, HLEG, LabCR, and IntCLIP using the same matched CLIP ViT-L/14 features and split. Specifically, FDIL is higher on both curves, and the accompanying table reports ROC-AUC, PR-AUC, mAP, confidence intervals, and paired tests. This keeps the visual comparison and the numerical evidence under the same controlled setting.
\revloc{Sec.~\ref{sec:experiments} (Calibration and Threshold-free Evaluation), Fig.~\ref{fig:roc_pr}, and Table~\ref{tab:threshold_free}}
\revshot{roc_pr}}

\cresp{A new section called ``Discussion'' could be added. The authors can discuss the main findings and also the important advantages and disadvantages of the proposed model.}
{Following this suggestion, we have added a Discussion that summarizes the main findings and advantages and states the observed failure cases and generalization limits, including the \textit{CreativeUnique} regression and the scope of the EMOTIC evidence.
\revloc{Sec.~\ref{sec:discussion}}
\revshot{discussion}}

\cresp{It would add more value to the paper if more theoretical analysis were conducted, e.g., from information-theoretical or feature selection perspectives.}
{We have added an information-theoretic excess-risk decomposition that separates label-side conditional entropy from feature-side inter-class aliasing. We also cast SLR-C as local feature selection and reranking: the candidate-recall diagnostic shows that most positives are already retrievable within the Top-10 set, so the remaining challenge is primarily to separate neighboring intents. The resulting predictions are evaluated by the component, negative-control, and unified-versus-decoupled experiments; the detailed control results are given in our response to Reviewer~2's novelty concern rather than repeated here.
\revloc{Sec.~\ref{sec:decoupling_principle}, Eq.~\eqref{eq:risk_decomp}, and Sec.~\ref{sec:experiments}, Tables~\ref{tab:unified_decoupled}, \ref{tab:negative_controls}, and~\ref{tab:threshold_free}}
\revshot{theory}}

\cresp{A comparative analysis of model parameters, FLOPs, or inference times is necessary.}
{We now quantify the deployment cost with a same-environment comparison of trainable parameters, FLOPs, and per-image latency for FDIL and the feature-level baselines, with all heads evaluated over the shared frozen CLIP features. Specifically, FDIL uses $2.41$M parameters, $4.93$M FLOPs, and $0.47$ms per image; the table contains the full method-by-method comparison. IntentMLM is identified separately as an MLLM pipeline whose cost is not directly comparable to a feature-level head.
\revloc{Sec.~\ref{sec:experiments} (Efficiency Comparison), Table~\ref{tab:efficiency_comparison}}
\revshot{efficiency}}
