% ================================
% === Reviewer 5 ===
% ================================
\reviewerblock{Reviewer~5}{%
We thank the reviewer for the positive assessment of the decoupling perspective and the module design, and for the precise minor corrections.}

\cresp{Figure~2 caption inconsistency: The caption refers to ``(Top) UTD'' and ``(Bottom) SLR-C'', but the figure shows ``A'' as the top and ``B'' as the bottom. Please revise the caption or the figure to ensure consistent correspondence.}
{Thank you for identifying this inconsistency. To satisfy the page limit, we shortened the Figure~2 caption and removed the former ``Top/Bottom'' panel description. The revised caption now summarizes the functions of SLR-C and UTD without making the inconsistent correspondence, so this issue no longer appears in the manuscript.
\revloc{Fig.~\ref{fig:fdil_framework} caption}
\revshot{framework_caption}}

\cresp{Missing experimental settings in Table~5:: Two additional configurations should be included: (i) Lexical + Scenario, and (ii) Canonical + Scenario.}
{We have added both missing pairwise configurations for a complete ablation: Lexical+Scenario and Canonical+Scenario under the same $K{=}10$ reranking-only protocol. Specifically, both improve over their single-source components; Canonical+Scenario is the strongest pair on Hard-F1, while the three-source ensemble remains best on aggregate Avg-F1. The table reports the complete metric values.
\revloc{Sec.~\ref{sec:experiments} (Heterogeneous Priors in SLR-C), Table~\ref{tab:prior_ablation}}
\revshot{prior_ablation}}

\cresp{Unexplained anomaly in Table~6: Provide a justification for why $K=5$ $K=28$ yield better performance than $K=10$.}
{We revisited this anomaly by rerunning $K\in\{5,10,28\}$ over three seeds under the final protocol. The revised sweep identifies \res{$K=10$} as the best setting, so the earlier single-seed dip is not used in our conclusion. Specifically, $K=5$ is slightly too restrictive and can exclude plausible positive labels, whereas $K=28$ admits all classes and adds distractor noise without improving semantic coverage. A moderate $K=10$ best balances candidate recall and local separability. We now use $K=10$ throughout and have added this analysis directly to the sensitivity experiment accompanying the complete metrics table.
\revloc{Sec.~\ref{sec:experiments} (Sensitivity to Candidate Size $K$), Table~\ref{tab:k_sensitivity_test}}
\revshot{k_sensitivity}}

\cresp{Typo in Appendix tables: In Table~B.1 and Table~B.2 (page~35), ``FDIR'' should be corrected to ``FDIL''.}
{Thank you for identifying this typo. To meet the page limit, we removed the former Appendix Tables~B.1--B.2 from the revised manuscript, so the reported instance no longer appears. We also performed a manuscript-wide terminology check and confirmed that every remaining occurrence uses the correct acronym ``FDIL''.
\revloc{Manuscript-wide terminology; former Appendix Tables~B.1--B.2}
\revshot{terminology}}
