\section{Related Work}
\label{sec:related_work}

\subsection{Image Intent Recognition}
Intent recognition seeks to decode the underlying purposes of humans. 
Early studies mainly focus on specific domains such as political persuasion \cite{joo_visual_2014, joo_automated_2015, huang_inferring_2016}, where intent is inferred from facial expressions, body language, and scene context. 
More recently, the release of the Intentonomy dataset \cite{jia_intentonomy_2021} has facilitated the study of general-purpose visual intention understanding in unconstrained social media scenarios.

Existing methods for visual intention recognition seek to solve the problem from different perspectives. 
Methods such as PIP-Net \cite{wang_prototype-based_2023} concentrate on feature and representation learning, which introduce prototype-based representations to capture to capture diverse visual patterns and improve interpretability. 
Considering the hierarchical nature of intent labels, approaches like CPAD \cite{ye_cross-modality_2023} and HLEG \cite{shi_learnable_2023} exploit label hierarchies to improve multi-level semantic alignment and mitigate ambiguity.  
LabCR \cite{shi_label-aware_2024} addresses label noise and shifting through calibration strategies, while recent works such as IntCLIP \cite{leonardis_synergy_2025} and MLMIntent \cite{wang_unlocking_2026} leverage vision-language models and textual priors to align visual features with semantic intent descriptions.

Despite these advances, most existing methods address ambiguity from a single perspective, such as feature representation, label structure, or external knowledge. 
They implicitly assume that ambiguity should be handled uniformly within a model. 
However, in visual intention recognition, ambiguity arises from different sources and may affect different stages of the recognition process in distinct ways.

\subsection{Learning under Ambiguous or Uncertain Supervision}

Learning under ambiguous or noisy supervision has been extensively studied in multi-label recognition and subjective understanding tasks, where label reliability is often compromised by annotation noise, semantic overlap, or annotator disagreement. 
Existing approaches typically address this issue through three main directions.

First, noise-robust learning methods mitigate unreliable supervision through robust objectives, hard-example handling, or noise-aware optimization \cite{lin_focal_2020, shrivastava_training_2016, ye_collaborative_2022, ridnik_asymmetric_2021}. 
These methods reduce the adverse impact of corrupted or hard labels, but they usually treat uncertainty as a generic optimization issue rather than modeling its semantic origin.

Second, hierarchical or structure-aware methods improve learning under ambiguous labels by exploiting label relations or taxonomy constraints \cite{zhang_use_2022, shi_learnable_2023}. 
Such methods are especially relevant to intention recognition, where semantically nearby labels often exhibit overlapping decision boundaries. 
However, they mainly focus on label dependency modeling, instead of explicitly using annotator disagreement to regulate supervision strength.

Third, knowledge distillation and vision-language supervision have been adopted to inject additional semantic guidance from teacher models or pretrained language priors \cite{touvron_training_2021, radford_learning_2021}. 
Knowledge distillation transfers structured knowledge from stronger teacher models, while vision-language models leverage textual descriptions as semantic anchors to enhance visual representations. 
However, these approaches typically assume that the provided semantic guidance is uniformly reliable across samples, which may not hold in subjective tasks with inherent ambiguity.

In visual intention understanding, ambiguity is intrinsic: the same image may support multiple plausible interpretations, and category boundaries are often fuzzy \cite{jia_intentonomy_2021, shi_learnable_2023}. 
Prior work has recognized the importance of hierarchical relations and label ambiguity in this task, but existing methods still largely rely on fixed supervision or globally applied semantic guidance. 
By contrast, our approach explicitly leverages annotator disagreement as an uncertainty signal, and adaptively regulates semantic guidance according to supervision ambiguity. 
This results in a more robust learning process under subjective and ambiguous supervision.