\section{Method}
\label{sec:method}

\subsection{Overview}

To address the dual challenges of supervisory ambiguity and semantic ambiguity in visual intention recognition, we propose \textbf{SADIR} (\textbf{S}tage-\textbf{A}ware \textbf{D}isambiguation for \textbf{I}ntent \textbf{R}ecognition), which explicitly decouples training and inference. Instead of handling these two forms of ambiguity within a unified mechanism, we address them in a stage-specific manner.

SADIR consists of two key components:
\begin{itemize}
    \item \textbf{Uncertainty-guided Text Distillation (UTD)}, which introduces adaptive semantic guidance during training to improve robustness under ambiguous supervision;
    \item \textbf{Semantic Local Reranking and Calibration (SLR-C)}, which performs lightweight semantic refinement during inference to enhance discrimination among competing intents.
\end{itemize}

Given an input image, the model first produces visual intent scores using a strong visual backbone. UTD is applied during training to shape the representation under uncertain supervision, while SLR-C refines predictions at inference time.

\subsection{Problem Formulation}

Given an input image $x_i$, the goal is to predict a set of intent labels from a predefined label space $\mathcal{Y}=\{1,\dots,C\}$. Due to the subjective nature of visual intention recognition, annotations are provided as soft label vectors $\mathbf{y}^{\text{soft}} \in \{0, 0.33, 0.66, 1\}^C$, reflecting annotator agreement.

Existing approaches typically follow either hard binarization or soft regression paradigms, both of which have limitations under ambiguous supervision. To address this, we adopt a \textbf{decoupled utilization} of annotation signals.

Specifically, we derive binarized labels $\mathbf{y}^{\text{hard}}$ for classification supervision, while extracting an image-level agreement score $\omega_i \in (0,1]$ from $\mathbf{y}^{\text{soft}}$. Here, $\omega_i$ reflects annotator consensus, where higher values indicate more reliable supervision.

This agreement score is used to adaptively control the contribution of semantic guidance during training. The model outputs a logit vector $\mathbf{s}^{v} \in \mathbb{R}^C$, which is later refined and thresholded for final prediction.

\subsection{Uncertainty-guided Text Distillation (UTD)}

\subsubsection{Motivation}

Supervisory ambiguity arises when multiple plausible intents coexist for a given image. Direct optimization against hard labels may lead to overconfident predictions and unstable decision boundaries.

To alleviate this issue, we introduce semantic guidance from text space and adapt its influence based on the uncertainty of supervision.

\subsubsection{Text-derived Teacher}

For each training image, we construct structured textual descriptions that capture semantic cues relevant to intention understanding. 
These descriptions are generated using a vision-language model, guided by both the image content and intent semantics.

To enhance discriminative capability under semantic ambiguity, we further incorporate contrastive cues by identifying visually plausible but incorrect intents predicted by the current model. 
Specifically, we select high-confidence negative intents as confusion candidates and use them to guide the generation of discriminative textual descriptions. 
This encourages the resulting text to emphasize subtle distinctions among semantically adjacent intents.

The generated text aims to summarize three aspects:
(1) observable visual elements,
(2) contextual interpretation beyond direct perception, and
(3) discriminative cues for distinguishing confusing intents.

The resulting text $T(x)$ is encoded into a dense embedding space using a frozen text encoder:
\begin{equation}
\mathbf{e}_{\text{text}} = f_{\text{text}}(T(x)).
\end{equation}

A lightweight classifier then maps the embedding into a soft probability distribution:
\begin{equation}
\mathbf{P}^{\text{tea}} = \text{Softmax}(g_{\text{tea}}(\mathbf{e}_{\text{text}}) / \tau),
\end{equation}
where $\tau$ is the temperature parameter.

This design introduces contrastive semantic signals that improve the teacher's ability to distinguish semantically similar intents, while remaining independent of the inference procedure.

\subsubsection{Uncertainty-guided Distillation}

The visual backbone produces logits $\mathbf{s}^{v}$ for image $x$. We adopt Asymmetric Loss \cite{ridnik_asymmetric_2021} for classification:
\begin{equation}
\mathcal{L}_{\text{vis}} = \frac{1}{C} \sum_{c=1}^{C} \ell_{\text{ASL}}(s^{v}_c, y_c).
\end{equation}

To incorporate semantic guidance, we introduce an uncertainty-aware distillation loss:
\begin{equation}
\mathcal{L}_{\text{distill}} = \text{KL}(\sigma(\mathbf{s}^{v}) \,\|\, \mathbf{P}^{\text{tea}}),
\end{equation}
where $\sigma(\cdot)$ is the sigmoid function.

The agreement score $\omega_i$ is used to balance the two objectives:
\begin{equation}
\mathcal{L}_{\text{total}} = \omega_i \cdot \mathcal{L}_{\text{vis}} + (1 - \omega_i) \cdot \lambda \cdot \mathcal{L}_{\text{distill}}.
\end{equation}

This formulation allows the model to rely more on visual supervision when annotations are reliable, and to leverage semantic guidance when supervision is uncertain.

\subsection{Semantic Local Reranking and Calibration (SLR-C)}

While UTD improves representation robustness, semantic ambiguity still affects final predictions, especially among visually similar intents. To address this, we propose SLR-C for inference-time refinement.

\subsubsection{Semantic Prior Construction}

Intent semantics are inherently multi-granular and abstract, making a single textual representation insufficient. We therefore construct a heterogeneous semantic prior for each intent by combining three complementary textual sources: lexical intent phrases, canonical intent descriptions, and descriptive visual scenarios.
The lexical source provides concise intent name cues, the canonical source provides semantic definitions, and the scenario source grounds abstract intents into plausible visual situations.

Given an image $x$, we compute image-text similarities:
\begin{equation}
s^{\text{prior}}_c = \text{sim}(f_{\text{img}}(x), \mathbf{e}_c).
\end{equation}

For each source, we obtain a class-wise similarity score and normalize it independently. The final prior is then constructed by averaging the normalized scores from the three sources:
\begin{equation}
\tilde{s}^{\text{prior}}_c = \frac{1}{3}\left(\tilde{s}^{\text{lex}}_c + \tilde{s}^{\text{can}}_c + \tilde{s}^{\text{scen}}_c\right).
\end{equation}

\subsubsection{Semantic Local Reranking}

We first select a compact candidate set:
\begin{equation}
\mathcal{C}_K(x) = \text{TopK}(\mathbf{s}^{v}, K).
\end{equation}

The logits are refined via local adjustment:
\begin{equation}
\tilde{s}_c =
\begin{cases}
s^{v}_c + \alpha \cdot \tilde{s}^{\text{prior}}_c, & c \in \mathcal{C}_K(x), \\
s^{v}_c, & \text{otherwise}.
\end{cases}
\end{equation}

This localized refinement focuses on resolving ambiguity among competing intents without disturbing the global prediction structure.

\subsubsection{Class-wise Calibration}

To account for class imbalance and varying score distributions, we adopt class-wise thresholds:
\begin{equation}
\hat{y}_c = \mathbb{I}[\sigma(\tilde{s}_c) > t_c],
\end{equation}
where $t_c$ is determined on the validation set.

\subsection{Stage-Aware Ambiguity Modeling}

We interpret visual intention recognition as a stage-dependent ambiguity problem. Supervisory ambiguity primarily affects representation learning, while semantic ambiguity affects decision boundaries.

SADIR addresses these two aspects in a stage-aware manner: UTD improves robustness during training through uncertainty-guided semantic distillation, while SLR-C enhances discrimination during inference via localized semantic refinement.

This stage-aware design enables effective handling of ambiguity without introducing heavy architectural complexity.
