\section{Related Work}
\label{sec:related_work}

\revised{Early visual intent studies target specific domains such as persuasive or political imagery~\cite{joo_visual_2014,joo_automated_2015,huang_inferring_2016}, while the Intentonomy benchmark~\cite{jia_intentonomy_2021} enables general-purpose intent understanding in unconstrained social media. Recent Intentonomy methods address the task from complementary angles: PIP-Net~\cite{wang_prototype-based_2023} uses prototype-based representations for interpretable intent perception; CPAD~\cite{ye_cross-modality_2023} and HLEG~\cite{shi_learnable_2023} exploit cross-modality or hierarchical label structures; LabCR~\cite{shi_label-aware_2024} models label calibration and relation preservation; and IntCLIP~\cite{leonardis_synergy_2025} and IntentMLM~\cite{wang_unlocking_2026} introduce CLIP/MLLM-based semantic knowledge. Robust-loss methods such as Focal Loss~\cite{lin_focal_2020}, OHEM~\cite{shrivastava_training_2016}, collaborative refining~\cite{ye_collaborative_2022}, and ASL~\cite{ridnik_asymmetric_2021} reduce the effect of difficult, noisy, or imbalanced samples through reweighting or selective optimization; however, they treat ambiguity mainly as an optimization issue rather than as a subjective semantic signal. Disagreement-aware learning instead preserves label diversity: Peterson et al.~\cite{peterson_human_2019} train on human label distributions, Davani et al.~\cite{mostafazadeh_davani_dealing_2022} model annotator-specific labels, DisCo~\cite{weerasooriya_disagreement_2023} jointly models item- and annotator-level distributions, and CrowdOpinion~\cite{weerasooriya_subjective_2023} recovers minority opinions. In contrast, FDIL explicitly treats annotator disagreement as an uncertainty signal and adaptively regulates semantic guidance according to supervision ambiguity, yielding more robust learning under subjective supervision.}

%% --- BEGIN condensed-out for 35-page limit (original detailed Related Work retained below) ---
\iffalse
\subsection{Image Intent Recognition}
Intent recognition seeks to decode the underlying purposes of humans. 
Early studies mainly focus on specific domains. \revised{Joo et al.~\cite{joo_visual_2014} infer communicative intent from persuasive images through facial and contextual cues, and subsequently relate facial-trait judgments to election outcomes~\cite{joo_automated_2015}; Huang and Kovashka~\cite{huang_inferring_2016} further infer visual persuasion by jointly modeling body language, scene setting, and deep features.}
More recently, the release of the Intentonomy dataset \cite{jia_intentonomy_2021} has facilitated the study of general-purpose visual intention understanding in unconstrained social media scenarios.

Existing methods tackle visual intention recognition from different perspectives. PIP-Net \cite{wang_prototype-based_2023} introduces prototype-based representations to capture diverse visual patterns and improve interpretability; CPAD \cite{ye_cross-modality_2023} and HLEG \cite{shi_learnable_2023} exploit label hierarchies for multi-level semantic alignment; LabCR \cite{shi_label-aware_2024} addresses label noise and shifting through calibration; and IntCLIP \cite{leonardis_synergy_2025} and IntentMLM \cite{wang_unlocking_2026} leverage vision-language models and textual priors to align visual features with intent descriptions. However, these methods address ambiguity from a single perspective and implicitly assume it should be handled uniformly, whereas in practice ambiguity arises from different sources and affects different stages of recognition in distinct ways.

\subsection{Learning under Ambiguous or Uncertain Supervision}

Learning under ambiguous or noisy supervision, where label reliability is compromised by annotation noise, semantic overlap, or annotator disagreement, has been addressed in three main directions.

First, noise-robust learning mitigates unreliable supervision through robust objectives and example reweighting: \revised{Focal Loss~\cite{lin_focal_2020} down-weights easy examples, online hard-example mining~\cite{shrivastava_training_2016} selects difficult samples during detector training, collaborative refining~\cite{ye_collaborative_2022} suppresses label noise in person re-identification, and the asymmetric loss~\cite{ridnik_asymmetric_2021} re-balances positive and negative gradients for multi-label classification.} These methods treat uncertainty as a generic optimization issue rather than modeling its semantic origin. A related line instead preserves disagreement as a learning signal: \revised{Peterson et al.~\cite{peterson_human_2019} train on human label distributions for robustness, Davani et al.~\cite{mostafazadeh_davani_dealing_2022} model individual annotators rather than the majority vote, DisCo~\cite{weerasooriya_disagreement_2023} jointly models item- and annotator-level label distributions, and CrowdOpinion~\cite{weerasooriya_subjective_2023} recovers meaningful minority opinions.} This is especially relevant to intention recognition, where multiple interpretations are often simultaneously plausible.

Second, hierarchical or structure-aware methods exploit label relations or taxonomy constraints \cite{zhang_use_2022, shi_learnable_2023} to handle semantically nearby labels with overlapping boundaries, but focus on label dependency rather than using disagreement to regulate supervision strength.

Third, knowledge distillation and vision-language supervision inject semantic guidance from teacher models or pretrained language priors \cite{touvron_training_2021, radford_learning_2021}, transferring structured knowledge or using textual descriptions as semantic anchors; however, they assume this guidance is uniformly reliable across samples, which may not hold under inherent ambiguity.

By contrast, our approach explicitly leverages annotator disagreement as an uncertainty signal and adaptively regulates semantic guidance according to supervision ambiguity, yielding more robust learning under subjective supervision.
\fi
%% --- END condensed-out for 35-page limit ---